Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2015-05-30 structured.tex
2015.05.30 (Vassilis) Changes added based on Hüseyin's comments. Mainly minor re-wording and fixing typos.

In terms of our $3$ stated criteria, the approach fares as follows:

@@

A noted Mithos limitation is that the protocol cannot effectively
handle dynamic arrivals and departures from the network~\cite{RGJZ2004}. 
\cite{CCRK2004}~points out extensive probing for Mithos to determine 
distances and less accurate proximity results compared to other
virtual coordinate systems. The last issue  has been also attributed
by~\cite{WSS2005} to the iterative neighbor selection process that is 
reportedly prone to premature termination at local minima.
%%
\cite{cox_vivaldi_2004} argues that elements of the protocol's
synthetic ID generation require a centralized implementation 
affecting Mithos' scalability.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In regard to the $3$ criteria, \emph{Mithos} stands as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% It has very good results in reducing communication latency among peers but its
% efficiency is being constrained by the fact that the proximity is calculated
% at bootstrap, assigning virtual coordinates that are used throughout of a
% nodes life cycle and thus the overlay is not adapting to a high-churn
% environment.
medium &
% Incremental probing is applied in quadrants and in two-hop distance at each
% step that does not generate a lot o control overhead.
medium &
% It is constrained by the fact that the protocol does not respond to dynamic
% peer arrivals and departures
low
\end{tabular}
}
\end{center}

@@

\emph{LAPTOP}'s behavior for the $3$ criteria stands as follows:

@@

When joining the network, each node
receives a $64$-bit \emph{ID}. 
It comprises of a random part and the IPv6 prefix which 
is obtained by hashing the node's IP address and port
number; this combination is deployed in an effort  
to preserve the uniformity of node assignment. 
The prefix in question is based on IPv6's global routing prefix.
As IANA\footnote{\url{http://www.iana.org/}}  
assigns address blocks to regions (RIRs) which are further
distributed at national level (NIRs), the first $32$-bits of the address
can estimate the region or the country the node is connected in. 
Thus, CHOP6's IPv6 prefix ID is used for proximity estimation
if no RTT information is available during routing. 
%%
%
%\paragraph{}
% There are three cases in which a node chooses the next hop according to
% information it possesses.
%\begin{itemize}
% \item When no information about RTT is available candidate next hops, the
%  sender just selects a node in the finger table entry which shares the longest
% prefix with the destination.
% \item There is another possibility that after some communication with other
%  nodes, the source node should know of the RTTs to some of the nodes in finger
% table entry. In this case, the source node chooses the one with the smallest
% RTT with probability $p$. If there is a node with no measured RTT then the
% sender can select such a node with probability $1 - p$
% \item In this last case the source node has already communicated with all
%  candidate node in the finger table entry. Thus, the node selects the node
% whose RTT is the smallest with probability $p$, the one with the second
% smallest with probability $q$ and so on, where $0 \leq \ldots < r < q < p < 1$
% and $p+q+r+\ldots \leq 1$
%\end{itemize}
%
With respect to the $3$ qualitative criteria, \emph{CHOP6} stands as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Simple approach that exploits proximity but it is coarse grained and thus not
% efficient.
low &
% Its static nature imposes low overhead to the network
low &
% IPv6 is still not widely applied and supported.
medium
\end{tabular}
}
\end{center}

@@

%%AD a better connection with the mismatch problem has to be made... 
%%		.. in the form of 1-2 lines (or short statements). 
%%
%% VM Added some extra lines and restructured the text
\emph{T2MC}'s~\cite{SLCGZ2008} objective is to diminish the redundant multiple
messages that may unnecessarily cross autonomous system (\emph{AS}) boundaries.
The latter represent links that are the costliest to ISPs and hence, 
minimizing topology mismatches  greatly affects pertinent operational expenditure.
%%
To achieve that, each node exploits the stable structure of
the Internet routers to help solve a clustering problem. 
The algorithm uses {\sl traceroute}-logs to detect inter-\emph{AS} boundaries 
and clusters peers to either nearby or 
far-away groups by applying a customized $k$-mean classification ($k=2$).
%%
When joining the network, the peer randomly picks an IP and uses
traceroute and selects the routers with
minimum and maximum latencies in the above {\sl traceroute} path 
as initial centroids for the ``nearby'' and ``far-away'' sets respectively.
Following an iterative procedure, the node comes up with the two sets 
demonstrating minimum intra-set latency variance.
At this stage, routers in the ``far-away'' set are not considered critical.
%%
%%AD skipping the details???? JUST CHECK
% Skipping the details of the algorithm, it picks the routers with
% minimum and maximum latencies in the above traceroute path as initial centroids
% for the ``near'' and ``remote'' sets respectively and after some iterations it
% ends up with two sets having the minimum intra-set latency variance. 
%%
% The router from the ``far-away'' set that presents the closest distance 
% from the peer in terms of hops, is designated as beacon.
% Any node portraying distance longer that the above is classified 
% as 
% 
% to the peer, in terms of hops, is designated as a beacon.
% Any node whose distance is measured to be longer that the distance 
% between 
% 
% The router with the minimum hops in the ``remote'' set is set up as the hop
% threshold and anything above that is classified as the ``remote'' router cluster
% for that peer. 
% Similarly, 
In contrast, the router from the ``near'' set
having the maximum number of hops 
from the just-joining peer is key as it denotes an edge--gateway. 
Peers that share the same edge--gateway or other members in the ``nearby'' 
router cluster are considered intra-\emph{AS} neighbors.
%%AD the final phrase for this is missing! So what??? What is achieved? 1 sentence.

@@

The \emph{ID} of an incoming node is computed based on 
both the identifier of the bootstrap node 
and the bootstrap's predecessor;
this is  done in a way that the \emph{ID} of the 
incoming peer is positioned \emph{between} the two identifiers mentioned above. 
After joining, a \emph{finger} table is created ala \emph{Chord}. 
Moreover, an \emph{adjacency table} maintains 
information about the $k$ closest--known peers. 
%%
The routing decision takes place in a way reminiscent to \emph{Chord}
and uses both adjacency and finger tables.
%%
\emph{AChord} behavior has as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% good routing efficiency but no adaptation to high churn environments.
medium &
% Relieves the system from joining overhead and topology aware ID construction
% but doubles the information needed to be stored and maintained by the
% introduction of adjacency table
medium &
% assumes the operation of ideal underlying any-cast mechanism (meaning that
% outside messages are always routed to the nearest node which in the current
% Internet, this is not always feasible. Moreover, is designed on the basis of
% IPv6 protocol which is not widely operational yet.
medium
\end{tabular}
}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{DHT-PNS}
A proximity neighbor selection scheme functioning atop 
the \emph{Chord DHT} is introduced in~\cite{DLTZZ2006}.
Its key objective was to group physically-close nodes as 
neighbors in the \emph{DHT} table while exploiting proximity information.
To detect proximity, virtual network coordinates for 
peers are used with the help of the Vivaldi protocol~\cite{cox_vivaldi_2004}. 
The virtual coordinates are then used to help map nodes to 
the identifier--space of the \emph{DHT}.
%%
The space in question is partitioned using a 
\emph{concentric circle clustering scheme} where successive cycles of
radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. 
The formed annuluses are then divided into 
$2\chi-1$ \emph{sectors}, where $\chi$ denotes the
sequence number of the annulus starting from $\chi = 1$ 
for the center cycle. 
In this way, 
\cite{DLTZZ2006} prove that each sector
This characteristic favors a more load--balanced clustering operation
should we assume uniform node distribution.
Every sector in this $2d$ space is mapped to 
a unique \emph{region} in the \emph{DHT} space forming a
multi-layer node identifier space.  
Consequently, nodes from the same
sector are mapped to the same region as well, 
preserving their proximity relationship.

Individual elements are uniquely mapped to the identifier--space,
allowing logarithmic lookup operations with high 
probability on \emph{Chord}. 
To this end, 
a node can obtain its \emph{DHT} key and its region key, 
in a fully distributed manner, just by applying a consistent hash function. 
The node can then communicate with the region's
master node called \emph{Cluster Node (CN)} which 
is responsible for clustering 
the nodes belonging in the same sector or 
region with that of the node at hand. 
%%
The node registers to its corresponding region  and publishes
its information, through the special peer \emph{CN};
the peer can also ask \emph{CN} for other
nodes that have previously joined the region to enhance 
the peer's neighbor set for future routing table optimization.
%
If no other nodes are found within the region, 
the search is expanded to neighboring regions.
%
%% VM rephrase the one bellow as above in a super compact form
%%AD the one below is not a parse-able sentence!!! What does it say???
%%%		restrict the length of your sentences to say at most 2.5 lines!
%Even in the case when no
%neighbor is detected in the current region, the search is expanded to adjacent
%regions and towards an upper layer identifier space until one is found or the
%first layer reached. Simulations suggest that the scheme has good convergence
%property of the delay distribution and good optimizing performance.
%
%
%%AD up to here the text in this second paragraph for Chord DHT remains 
%%%%%		... rather cryptic - does need work and better connection 
%%%%%		with the 1st paragraph...
%%
Even though scaling appeared promising, only limited 
simulations were carried out with configurations involving up to 2,048 nodes.
Moreover the approach assumes uniform node distribution which
is not always the case, especially in a synthetic coordinate system.
%%
\emph{Chord DHT} fares as follows in reference to the stated $3$ criteria:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
medium &
%
medium &
%
medium
\end{tabular}
}
\end{center}

@@

the $2$-dimensional is converted to $1$-dimensional space. 

@@

\emph{Quasi-Chord} qualitative behavior has as follows:

@@

Each peer in the overlay has a unique \emph{ID} and and an IP address to
characterize it. 
To publish itself, the peer hashes a fixed-length prefix of 
its IP to obtain a key and stores the resulting \emph{ID} and IP 
in the DHT. 
Newly joining nodes with the same {\sl IP} prefix 
can, thus, query the \emph{DHT} and readily identify all 
neighbors with the same prefix. 
Nodes periodically update their entries in the \emph{DHT} 
by removing those who leave or have timed-out.
%%

@@

A new-coming peer then computes distances from those landmarks,
sorts them, and assigns itself an \emph{ID} on this 
landmark ordering as follows:
the peer adopts sets the prefix of its \emph{ID} to be equal to that 
of its closest landmark.
The length of this prefix is 
$l_{prefix}=|log_b k|$, where $b$ is the \emph{ID}--base 
and $k$ the number of landmarks. 
%%
The remaining of bits of the peer-\emph{ID} are assigned either 
randomly or computed with additional proximity optimizations.
In this way, two desirable features are attained:
\begin{inparaenum}[\itshape i\upshape)]
  \item landmark assignment is dynamic, and
  \item the set of a node leaf may reference peers that are 
	physically close by.
\end{inparaenum}
The latter in particular helps reduce average routing cost(s)
for the leaf-set of a node is used to efficiently determine the 
last routing step.

\emph{CNPA} is proposed for network setups for which a lightweight
bootstrap phase is required.
It delegates the task of identifying the physically
close--node to \emph{Pastry}; this node is used by the 
incoming peer to get bootstrapped.
The new-comer takes the bootstrapping nodes's \emph{ID} prefix 
and fills the remaining bits of its key using the \emph{RLM} technique.
%%
\emph{RLM} gets evaluated as presenting good proximity in a load-balanced 
environments. Moreover, it is adaptable to high-churn settings.
Unfortunately, \emph{RLM} bootstrap overhead is substantial and can 
effect the set of landmarks~\cite{WZS2004}.
\emph{CNPA} on the other hand, incurs less overhead but this come 
at the expense of being more coarse-grained as it is based on prefixes
to determine proximity during.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \emph{RLM} achieves good proximity evaluation in a load balanced and adaptable to
% high churn environments fashion. Unfortunately it comes with more bootstrap
% overhead that can especially effect the set of temporary landmarks. On the other
% hand, \emph{CNPA}'s less incurred overhead comes at the expense of being more
% coarse-grained.
%%%
%%
% \emph{DynaMO} tries to capitalize on the fact that in Pastry each node's table consists
% of rows equal to the number of digits in the overlay ID and columns equal to
% the ID's base. As we go down the table rows, the matching prefix between the
% current node's ID and the row's entries increases by one. Thus, the leaf set
% contains the closest nodes in the ID space. Additionally, as the prefix match
% increases by one, the result is exponentially fewer candidates that can fill the
% tables entries as the row increases. This leads to the observation
%  In this context, two approaches are considered namely
% \emph{Random Landmarking (RLM)} and \emph{Closest Neighbor Prefix Assignment
% (CNPA)}.
%
% TODO: SOME DISCUSSION
%
%Both algorithms, are protected against the formation of physical landmark
%  clusters or imbalanced ID distribution\footnote{More common during the
% initial formation of an overlay network} , by introducing the \emph{landmark
% gravitation range} as a threshold over which landmark keys are reassigned (for
% the RLM approach) or unutilized ID prefix ranges are detected and used (for
% the CPNA scheme) in order to balance the distribution of regions in the
% overlay.
%
\emph{DynoMO}'s elements fare as follows regarding our $3$ criteria:
\begin{center}
{\footnotesize
\begin{tabular}{rccc}
\multicolumn{1}{r}{} &
\multicolumn{1}{c}{\emph{Efficiency}} &
\multicolumn{1}{c}{\emph{Overhead}} &
\multicolumn{1}{c}{\emph{Scalability}}
\\
\cline{2-4}
\emph{RLM} &
%%
%% TAKEN FROM
%% Efficient service discovery mechanism for wireless sensor networks
%%
%%
% common prefix of the overlay ID with the closest landmark. The
% RLM utilizes the overlay lookup capabilities by reducing traffic
% overhead within an overlay cluster. However, it generates much
% control traffic for periodically measuring distances to all landmarks.
% Moreover, the hop stretch overhead still increases when
% performing overlay routing among overlay clusters.
medium &
%
medium &
% Dynamic landmarking approach helps the scalability of the algorithm. The
% authors claim that the temporary nature of RLM's landmark nodes does not
% prevent them to maintain a good ratio in the presence of high landmark
% digression rates.
high \\
\emph{CNPA} &
%
medium &
% % The paper suggests that this approach generates significantly lower overhead
% than the vanilla Pastry algorithm
low &
% The queries are directed through fast nodes so it is important such nodes
% be part of the network
medium \\
\end{tabular}
}
\end{center}

@@

\emph{3DO} stores objects in the same $3$-$d$ space. For each
file $f$, a key $k$ of the form of an $x$, $y$ and $z$ tuple
is hashed out of the file's unique \emph{ID}~(e.g., filename).
The node whose \emph{LID} is closer to $k$ stores the
file itself.
%%
To retrieve the file, the requesting node calculates $f$'s $k$
and issues a query message. 
The message is a associated with a \emph{TTL} value. 
Each peer forwards that query to the node,
in its routing table, that is closest to $k$ until either the peer
that stores the file is located or the query's lifetime expires.

@@

In what follows, we place the surveyed \emph{structured} \p-methods in 
the three above categories formulated based on 
utilized proximity and geographic aspects.
Table~\ref{structured:table} offers a comparative view
of the algorithms in a way reminiscent to that used in Table~\ref{unstructured:table}.
Table~\ref{structured:table} outlines key properties for all surveyed approach 
that helps differentiate and better position competing techniques.
Clearly, we do not further elaborate topics that we have included and discussed 
at length in Section~\ref{section:unstructured}; these aspects entail 
topology adaptation, landmarking, and caching/replication.
Finally, Figures~\ref{structured:plot:efficiency}, \ref{structured:plot:overhead}
and \ref{structured:plot:scalability} help with the pictorial comparison
of the algorithms' bahavior with respect to the three effectiveness
criteria we defined in subsection~\ref{background:motivation}.

@@

%% VM: I reworded the above as below.
This category consists of methods that try to map the overlay id space on to
the physical one, as depicted in Figure~\ref{figure:geographic-layout}.
Routing tables of involved peers, are adjusted and maintained by exploiting
proximity information that describes the overall (geographic) positioning of
the participants. This comes in contrast to proximity routing and proximity neighbor
selection (discussed later on) where decisions are taken based on and effecting
a peer's nearby environment.

\begin{figure}[ht]
\centering
  \includegraphics[scale=0.44]{img/pdf/geographic-layout.pdf}
\caption{Successive nodes in the overlay circle can be chosen to be from the
same or near autonomous systems (\emph{AS}).}
\label{figure:geographic-layout}
\end{figure}
%%
A number of the reviewed algorithms including 
\emph{Global Soft-State},
\emph{Mithos} and \emph{LAPTOP} resort to clustering close-by nodes in
order to reduce the network's average delay.
%%
\input{structured-table}
%%
\input{structured-plot}
%%
Landmark servers and \emph{RTT} measurements are two popular methods, which
can also be used in conjunction, to discover peers in physical proximity.
However, as already discussed, these
methods do not always yield reliable estimates for node positions over the
Internet. 
Landmark servers are not self-organizing and present maintenance
overhead. Moreover, to serve a structured \p--network with potentially 
tens of thousands or more 
% millions 
peers, multiple such servers are required to be distributed over the world, 
which is a difficult proposition per se.
\emph{RTT}s can measure the delay between peers, but their use constitutes a
greedy method that can result in sub-optimal overlay topologies;
the latter is especially true if nearby 
nodes happen to be connected through low bandwidth connections.

>>>>

In terms of our three stated criteria, the approach fares as follows:

@@

A noted Mithos limitation is that the protocol cannot effectively
handle dynamic arrivals and departures from the network~\cite{RGJZ2004}. 
\cite{CCRK2004}~points out extensive probing for Mithos to determine 
distances and less accurate proximity results compared to other
virtual coordinate systems. The last issue  has been also attributed
by~\cite{WSS2005} to the iterative neighbor selection process that is 
reportedly prone to premature termination at local minima.
%%
\cite{cox_vivaldi_2004} argues that elements of the protocol's
synthetic ID generation require a centralized implementation 
affecting Mithos' scalability.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In regard to the three criteria, \emph{Mithos} stands as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% It has very good results in reducing communication latency among peers but its
% efficiency is being constrained by the fact that the proximity is calculated
% at bootstrap, assigning virtual coordinates that are used throughout of a
% nodes life cycle and thus the overlay is not adapting to a high-churn
% environment.
medium &
% Incremental probing is applied in quadrants and in two-hop distance at each
% step that does not generate a lot o control overhead.
medium &
% It is constrained by the fact that the protocol does not respond to dynamic
% peer arrivals and departures
low
\end{tabular}
}
\end{center}

@@

\emph{LAPTOP}'s behavior for the three criteria stands as follows:

@@

When joining the network, each node
receives a $64$-bit \emph{ID}. 
It comprises of a random part and the IPv6 prefix which 
is obtained by hashing the node's IP address and port
number; this combination is deployed in an effort  
to preserve the uniformity of node assignment. 
The prefix in question is based on IPv6's global routing prefix.
As IANA\footnote{\url{http://www.iana.org/}}  
assigns address blocks to regions (RIRs) which are further
distributed at national level (NIRs), the first $32$-bits of the address
can estimate the region or the country the node is connected in. 
Thus, CHOP6's IPv6 prefix ID is used for proximity estimation
if no RTT information is available during routing. 
%%
%
%\paragraph{}
% There are three cases in which a node chooses the next hop according to
% information it possesses.
%\begin{itemize}
% \item When no information about RTT is available candidate next hops, the
%  sender just selects a node in the finger table entry which shares the longest
% prefix with the destination.
% \item There is another possibility that after some communication with other
%  nodes, the source node should know of the RTTs to some of the nodes in finger
% table entry. In this case, the source node chooses the one with the smallest
% RTT with probability $p$. If there is a node with no measured RTT then the
% sender can select such a node with probability $1 - p$
% \item In this last case the source node has already communicated with all
%  candidate node in the finger table entry. Thus, the node selects the node
% whose RTT is the smallest with probability $p$, the one with the second
% smallest with probability $q$ and so on, where $0 \leq \ldots < r < q < p < 1$
% and $p+q+r+\ldots \leq 1$
%\end{itemize}
%
With respect to the three qualitative criteria, \emph{CHOP6} stands as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Simple approach that exploits proximity but it is coarse grained and thus not
% efficient.
low &
% Its static nature imposes low overhead to the network
low &
% IPv6 is still not widely applied and supported.
medium
\end{tabular}
}
\end{center}

@@

%%AD a better connection with the mismatch problem has to be made... 
%%		.. in the form of 1-2 lines (or short statements). 
%%
%% VM Added some extra lines and restructured the text
\emph{T2MC}'s~\cite{SLCGZ2008} objective is to diminish the redundant multiple
messages that may unnecessarily cross autonomous system (\emph{AS}) boundaries.
Inter-\emph{AS} links are the costliest to ISPs and hence, 
minimizing topology mismatches greatly affects pertinent operational expenditure.
%%
To achieve that, each node exploits the stable structure of
the Internet routers to help solve a clustering problem. 
The algorithm uses {\sl traceroute}-logs to detect inter-\emph{AS} boundaries 
and clusters peers to either nearby or 
far-away groups by applying a customized $k$-mean classification ($k=2$).
%%
When joining the network, the peer randomly picks an IP and uses
traceroute and selects the routers with
minimum and maximum latencies in the above {\sl traceroute} path 
as initial centroids for the ``nearby'' and ``far-away'' sets respectively.
Following an iterative procedure, the node comes up with the two sets 
demonstrating minimum intra-set latency variance.
At this stage, routers in the ``far-away'' set are not considered critical.
%%
%%AD skipping the details???? JUST CHECK
% Skipping the details of the algorithm, it picks the routers with
% minimum and maximum latencies in the above traceroute path as initial centroids
% for the ``near'' and ``remote'' sets respectively and after some iterations it
% ends up with two sets having the minimum intra-set latency variance. 
%%
% The router from the ``far-away'' set that presents the closest distance 
% from the peer in terms of hops, is designated as beacon.
% Any node portraying distance longer that the above is classified 
% as 
% 
% to the peer, in terms of hops, is designated as a beacon.
% Any node whose distance is measured to be longer that the distance 
% between 
% 
% The router with the minimum hops in the ``remote'' set is set up as the hop
% threshold and anything above that is classified as the ``remote'' router cluster
% for that peer. 
% Similarly, 
In contrast, the router from the ``near'' set
having the maximum number of hops 
from the just-joining peer is key as it denotes an edge--gateway. 
Peers that share the same edge--gateway or other members in the ``nearby'' 
router cluster are considered intra-\emph{AS} neighbors.
%%AD the final phrase for this is missing! So what??? What is achieved? 1 sentence.

@@

The \emph{ID} of an incoming node is computed based on 
both the identifier of the bootstrap node 
and the bootstrap's predecessor;
this is  done in a way that the \emph{ID} of the 
incoming peer is positioned \emph{between} the two identifiers mentioned above. 
After joining, a \emph{finger} table is created ala \emph{Chord}. 
Moreover, an \emph{adjacency table} maintains 
information about the $k$ closest--known peers. 
%%
The routing decision takes place in a way reminiscent to \emph{Chord}
and uses both adjacency and finger tables.
%%
\emph{AChord} behaves as follows:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% good routing efficiency but no adaptation to high churn environments.
medium &
% Relieves the system from joining overhead and topology aware ID construction
% but doubles the information needed to be stored and maintained by the
% introduction of adjacency table
medium &
% assumes the operation of ideal underlying any-cast mechanism (meaning that
% outside messages are always routed to the nearest node which in the current
% Internet, this is not always feasible. Moreover, is designed on the basis of
% IPv6 protocol which is not widely operational yet.
medium
\end{tabular}
}
\end{center}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{DHT-PNS}
A proximity neighbor selection scheme functioning atop 
the \emph{Chord DHT} is introduced in~\cite{DLTZZ2006}.
Its key objective was to group physically-close nodes as 
neighbors in the \emph{DHT} table while exploiting proximity information.
To detect proximity, virtual network coordinates for 
peers are used with the help of the Vivaldi protocol~\cite{cox_vivaldi_2004}. 
The virtual coordinates are then used to help map nodes to 
the identifier--space of the \emph{DHT}.
%%
The space in question is partitioned using a 
\emph{concentric circle clustering scheme} where successive cycles of
radiuses $\rho$, $2\rho$, $3\rho$ and so on, are constructed. 
The formed annuluses are then divided into 
$2\chi-1$ \emph{sectors}, where $\chi$ denotes the
sequence number of the annulus starting from $\chi = 1$ 
for the center cycle. 
In this way, 
\cite{DLTZZ2006} prove that for each sector
this characteristic favors a more load--balanced clustering operation
should we assume uniform node distribution.
Every sector in this $2d$ space is mapped to 
a unique \emph{region} in the \emph{DHT} space forming a
multi-layer node identifier space.  
Consequently, nodes from the same
sector are mapped to the same region as well, 
preserving their proximity relationship.

Individual elements are uniquely mapped to the identifier--space,
allowing logarithmic lookup operations with high 
probability on \emph{Chord}. 
To this end, 
a node can obtain its \emph{DHT} key and its region key, 
in a fully distributed manner, just by applying a consistent hash function. 
The node can then communicate with the region's
master node called \emph{Cluster Node (CN)} which 
is responsible for clustering 
the nodes belonging in the same sector or 
region with that of the node at hand. 
%%
The node registers to its corresponding region  and publishes
its information, through the special peer \emph{CN};
the peer can also ask \emph{CN} for other
nodes that have previously joined the region to enhance 
the peer's neighbor set for future routing table optimization.
%
If no other nodes are found within the region, 
the search is expanded to neighboring regions.
%
%% VM rephrase the one bellow as above in a super compact form
%%AD the one below is not a parse-able sentence!!! What does it say???
%%%		restrict the length of your sentences to say at most 2.5 lines!
%Even in the case when no
%neighbor is detected in the current region, the search is expanded to adjacent
%regions and towards an upper layer identifier space until one is found or the
%first layer reached. Simulations suggest that the scheme has good convergence
%property of the delay distribution and good optimizing performance.
%
%
%%AD up to here the text in this second paragraph for Chord DHT remains 
%%%%%		... rather cryptic - does need work and better connection 
%%%%%		with the 1st paragraph...
%%
Even though scaling appeared promising, only limited 
simulations were carried out with configurations involving up to 2,048 nodes.
Moreover the approach assumes uniform node distribution which
is not always the case, especially in a synthetic coordinate system.
%%
\emph{Chord DHT} fares as follows in reference to the stated three criteria:
\begin{center}
{\footnotesize
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
%
medium &
%
medium &
%
medium
\end{tabular}
}
\end{center}

@@

the $2$-dimensional space is converted to $1$-dimensional space. 

@@

\emph{Quasi-Chord}'s qualitative behavior is as follows:

@@

Each peer in the overlay has a unique \emph{ID} and an IP address to
characterize it. 
To publish itself, the peer hashes a fixed-length prefix of 
its IP to obtain a key and stores the resulting \emph{ID} and IP 
in the DHT. 
Newly joining nodes with the same {\sl IP} prefix 
can, thus, query the \emph{DHT} and readily identify all 
neighbors with the same prefix. 
Nodes periodically update their entries in the \emph{DHT} 
by removing those who leave or have timed-out.
%%

@@

A new-coming peer then computes distances from those landmarks,
sorts them, and assigns itself an \emph{ID} on this 
landmark ordering as follows:
the peer sets the prefix of its \emph{ID} to be equal to that 
of its closest landmark.
The length of this prefix is 
$l_{prefix}=|log_b k|$, where $b$ is the \emph{ID}--base 
and $k$ the number of landmarks. 
%%
The remaining of bits of the peer-\emph{ID} are assigned either 
randomly or computed with additional proximity optimizations.
In this way, two desirable features are attained:
\begin{inparaenum}[\itshape i\upshape)]
  \item landmark assignment is dynamic, and
  \item the set of a node leaf may reference peers that are 
	physically close by.
\end{inparaenum}
The latter in particular helps reduce average routing cost(s)
for the leaf-set of a node is used to efficiently determine the 
last routing step.

\emph{CNPA} is proposed for network setups for which a lightweight
bootstrap phase is required.
It delegates the task of identifying the physically
close--node to \emph{Pastry}; this node is used by the 
incoming peer to get bootstrapped.
The new-comer takes the bootstrapping nodes's \emph{ID} prefix 
and fills the remaining bits of its key using the \emph{RLM} technique.
%%
\emph{RLM} gets evaluated as presenting good proximity in a load-balanced 
environments. Moreover, it is adaptable to high-churn settings.
Unfortunately, \emph{RLM} bootstrap overhead is substantial and can 
effect the set of landmarks~\cite{WZS2004}.
\emph{CNPA} on the other hand, incurs less overhead but this comes 
at the expense of being more coarse-grained as it is based on prefixes
to determine proximity during bootstrap.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \emph{RLM} achieves good proximity evaluation in a load balanced and adaptable to
% high churn environments fashion. Unfortunately it comes with more bootstrap
% overhead that can especially effect the set of temporary landmarks. On the other
% hand, \emph{CNPA}'s less incurred overhead comes at the expense of being more
% coarse-grained.
%%%
%%
% \emph{DynaMO} tries to capitalize on the fact that in Pastry each node's table consists
% of rows equal to the number of digits in the overlay ID and columns equal to
% the ID's base. As we go down the table rows, the matching prefix between the
% current node's ID and the row's entries increases by one. Thus, the leaf set
% contains the closest nodes in the ID space. Additionally, as the prefix match
% increases by one, the result is exponentially fewer candidates that can fill the
% tables entries as the row increases. This leads to the observation
%  In this context, two approaches are considered namely
% \emph{Random Landmarking (RLM)} and \emph{Closest Neighbor Prefix Assignment
% (CNPA)}.
%
% TODO: SOME DISCUSSION
%
%Both algorithms, are protected against the formation of physical landmark
%  clusters or imbalanced ID distribution\footnote{More common during the
% initial formation of an overlay network} , by introducing the \emph{landmark
% gravitation range} as a threshold over which landmark keys are reassigned (for
% the RLM approach) or unutilized ID prefix ranges are detected and used (for
% the CPNA scheme) in order to balance the distribution of regions in the
% overlay.
%
\emph{DynoMO}'s elements fare as follows regarding our three criteria:
\begin{center}
{\footnotesize
\begin{tabular}{rccc}
\multicolumn{1}{r}{} &
\multicolumn{1}{c}{\emph{Efficiency}} &
\multicolumn{1}{c}{\emph{Overhead}} &
\multicolumn{1}{c}{\emph{Scalability}}
\\
\cline{2-4}
\emph{RLM} &
%%
%% TAKEN FROM
%% Efficient service discovery mechanism for wireless sensor networks
%%
%%
% common prefix of the overlay ID with the closest landmark. The
% RLM utilizes the overlay lookup capabilities by reducing traffic
% overhead within an overlay cluster. However, it generates much
% control traffic for periodically measuring distances to all landmarks.
% Moreover, the hop stretch overhead still increases when
% performing overlay routing among overlay clusters.
medium &
%
medium &
% Dynamic landmarking approach helps the scalability of the algorithm. The
% authors claim that the temporary nature of RLM's landmark nodes does not
% prevent them to maintain a good ratio in the presence of high landmark
% digression rates.
high \\
\emph{CNPA} &
%
medium &
% % The paper suggests that this approach generates significantly lower overhead
% than the vanilla Pastry algorithm
low &
% The queries are directed through fast nodes so it is important such nodes
% be part of the network
medium \\
\end{tabular}
}
\end{center}

@@

\emph{3DO} stores objects in the same $3$-$d$ space. For each
file $f$, a key $k$ of the form of an $x$, $y$ and $z$ tuple
is hashed out of the file's unique \emph{ID}~(e.g., filename).
The node whose \emph{LID} is closer to $k$ stores the
file itself.
%%
To retrieve the file, the requesting node calculates $f$'s $k$
and issues a query message. 
The message is associated with a \emph{TTL} value. 
Each peer forwards that query to the node,
in its routing table, that is closest to $k$ until either the peer
that stores the file is located or the query's lifetime expires.

@@

In what follows, we place the surveyed \emph{structured} \p-methods in 
the three above categories formulated based on 
utilized proximity and geographic aspects.
Table~\ref{structured:table} offers a comparative view
of the algorithms in a way reminiscent to that used in Table~\ref{unstructured:table}.
Table~\ref{structured:table} outlines key properties for all surveyed approach 
that helps differentiate and better position competing techniques.
Clearly, we do not further elaborate topics that we have included and discussed 
at length in Section~\ref{section:unstructured}; these aspects entail 
topology adaptation, landmarking, and caching/replication.
Finally, Figures~\ref{structured:plot:efficiency}, \ref{structured:plot:overhead}
and \ref{structured:plot:scalability} help with the pictorial comparison
of the algorithms' behavior with respect to the three effectiveness
criteria we defined in subsection~\ref{background:motivation}.

@@

%% VM: I reworded the above as below.
This category consists of methods that try to map the overlay id space on to
the physical one, as depicted in Figure~\ref{figure:geographic-layout}.
\begin{figure}[ht]
\centering
  \includegraphics[scale=0.44]{img/pdf/geographic-layout.pdf}
\caption{Successive nodes in the overlay circle can be chosen to be from the
same or near autonomous systems (\emph{AS}).}
\label{figure:geographic-layout}
\end{figure}
Routing tables of involved peers, are adjusted and maintained by exploiting
proximity information that describes the overall (geographic) positioning of
the participants. This comes in contrast to proximity routing and proximity neighbor
selection (discussed later on) where decisions are taken based on and effecting
a peer's nearby environment.

A number of the reviewed algorithms including 
\emph{Global Soft-State},
\emph{Mithos} and \emph{LAPTOP} resort to clustering close-by nodes in
order to reduce the network's average delay.
Landmark servers and \emph{RTT} measurements are two popular methods, which
can also be used in conjunction, to discover peers in physical proximity.
However, as already discussed, these
methods do not always yield reliable estimates for node positions over the
Internet. 
Landmark servers are not self-organizing and present maintenance
overhead. Moreover, to serve a structured \p--network with potentially 
tens of thousands or more 
% millions 
peers, multiple such servers are required to be distributed over the world, 
which is a difficult proposition per se.
\emph{RTT}s can measure the delay between peers, but their use constitutes a
greedy method that can result in sub-optimal overlay topologies;
the latter is especially true if nearby 
nodes happen to be connected through low bandwidth connections.
